---
title: "BE601 - Data Analytics I"
subtitle: "Seminar 2"
author: "Uyen Vu (uyen.vu@phdstudent.hhs.se)"
date: "13/08/2022"
output: word_document
monofont: 'consolas'
---

\newpage

# Test Exercise
# Discussion exercises
## 46 - page 

## 51

## E6

We define the events:

d1: item being defective

d2: item not being defective

c1: item being classified as defective

c2: item being classified as not_defective

```{r}
#5% of the produced items are defective
p_d1 <- 0.05
p_d2 <- 0.95 #complementary rule
#items that are free from defects are classified as correct with probability 0.98
p_c2_on_d2 <- 0.98
p_c1_on_d2 <- 1 - p_c2_on_d2 #complementary rule
p_c2_intersect_d2 <- p_c2_on_d2 * p_d2 #multiplication rule
p_c1_intersect_d2 <- p_c1_on_d2 * p_d2 #multiplication rule
#defective items are classified as defective with probability 0.95
p_c1_on_d1 <- 0.95
p_c2_on_d1 <- 1 - p_c1_on_d1 #complementary rule
p_c1_intersect_d1 <- p_c1_on_d1 * p_d1 #multiplication rule
p_c2_intersect_d1 <- p_c2_on_d1 * p_d1 #multiplication rule
# a) proportion of produced items classified as defective: p_c1
## Because c(c1,c2) and d(d1,d2) are independent; and each broken into mutually exclusive and collectively exhaustive:
p_c1 <- p_c1_intersect_d1 + p_c1_intersect_d2
p_c1
# b) Find the probability that an item classified as defective is defective: p_d1_on_c1
## Applying Bayes's theorm
p_d1_on_c1_nominator <- p_c1_on_d1 * p_d1
p_d1_on_c1_denominator <- p_c1_on_d1 * p_d1 + p_c1_on_d2 * p_d2
p_d1_on_c1 <- p_d1_on_c1_nominator/p_d1_on_c1_denominator
p_d1_on_c1
```

We can calculate the marginal probability of each event by using two-way table:

|            |       **D1**       |      **D2**      |      **Totals**      |
|------------|:------------------:|:----------------:|:--------------------:|
| **C1**     | 0.95*0.05=0.0475   | 0.02*0.95=0.019  | 0.0475+0.019=0.0665  |
| **C2**     | 0.05*0.05=0.0025   | 0.98*0.95=0.931  | 0.0025+0.0931=0.9335 |
| **Totals** | 0.0475+0.0025=0.05 | 0.019+0.931=0.95 |           1          |

#Additional recommended exercises
## 40
## 3.99

S1: Customer is student | S1: Customer is not student

V1: Customer eats vegetarian | V2: Customer does not eat vegetarian

```{r}
#35% of all customers order vegetarian meals and 50% of all customers are students.
p_v1 <- 0.35
p_v2 <- 1 - p_v1
p_s1 <- 0.5
p_s2 <- 1 - p_s1
#25% of all customers who are students order vegetarian meals
p_v1_on_s1 <- 0.25
p_v2_on_s1 <- 1 - p_v1_on_s1
#a) randomly chosen customer both is a student and eats vegetarian: p_s1_intersect_v1
p_s1_intersect_v1 <- p_v1_on_s1 * p_s1
p_s1_intersect_v1
#b) p_s1_on_v1
p_s1_on_v1 <- p_s1_intersect_v1 / p_v1
p_s1_on_v1
#c) p_s2_intersect_v2 (we can draw a 2-way table)
p_s1_intersect_v2 <- p_v2_on_s1 * p_s1
p_s2_intersect_v2 <- p_v2 - p_s1_intersect_v2
p_s2_intersect_v2
#d) independent events?
p_v1 * p_s1 == p_s1_intersect_v1
#e) mutually exclusive?
p_s1_intersect_v1 == 0
#f) collectively exhaustive?
p_s1 + p_v1 == 1
```

## 3.102

Define the events:

G1: worker is female | G2: worker is male

N1: worker works night-shift | N2: worker does not work night-shift

F1: worker favors the plan | F2: worker does not favor the plan

```{r}
#65% of all night_shift workers favored the plan
p_f1_on_n1 <- 0.65
#40% of all female workers favored the plan
p_f1_on_g1 <- 0.4
#50% of all employees are night-shift workers
p_n1 <- 0.5
p_n2 <- 1 - p_n1 #complementary
#30% of all employees are women
p_g1 <- 0.3
p_g2 <- 1 - p_g1
#20% of all night-shift workers are women
p_g1_on_n1 <- 0.2
#a) a randomly chosen employee is a woman in favor of the plan: p_g1_intersect_f1
p_g1_intersect_f1 <- p_f1_on_g1 * p_g1 #multiplication rule
p_g1_intersect_f1
#b) a randomly chosen employee is either a woman or a nigh-shift worker (or both): p_g1_intersect_n1 & p_g1_union_n1
p_g1_intersect_n1 <- p_g1_on_n1 * p_n1
p_g1_intersect_n1
p_g1_union_n1 <- p_g1 + p_n1 - p_g1_intersect_n1
p_g1_union_n1
#c) employee gender independent of night shift is worked?
p_g1_intersect_n1 == p_g1 * p_n1 #not independent
#d) female employee is a night-shift worker: p_n1_on_g1
p_n1_on_g1 <- p_g1_intersect_n1/p_g1
p_n1_on_g1
#50% of all male employees favor the plan
p_f1_on_g2 <- 0.5
#e) employee both does work the night shift and does not favor the plan: p_n2_intersect_f2
p_g2_intersect_f1 <- p_f1_on_g2 * p_g2
p_f1 <- p_g1_intersect_f1 + p_g2_intersect_f1
p_n1_intersect_f1 <- p_f1_on_n1 * p_n1
p_n2_intersect_f1 <- p_f1 - p_n1_intersect_f1
p_n2_intersect_f2 <- p_n2 - p_n2_intersect_f1
p_n2_intersect_f2
```

## 3.101

f1: employee is female | f2: employee is male
g: employee had graduate degree | u: employee had undergraduate degree | h: employee had high school degree

```{r}
#80% of the employees are men & 20% of the employees are women
p_f1 <- 0.2
p_f2 <- 0.8
#men: 10% had graduate, 30% had undergraduate, 60% had high school
p_g_on_f2 <- 0.1
p_u_on_f2 <- 0.3
p_h_on_f2 <- 0.6
#women: 15% had graduate, 40% had undergraduate, 45% had high school
p_g_on_f1 <- 0.15
p_u_on_f1 <- 0.4
p_h_on_f1 <- 0.45
#a) p_f2_intersect_h
p_f2_intersect_h <- p_h_on_f2 * p_f2
p_f2_intersect_h
#b) p_g
p_g_intersect_f1 <- p_g_on_f1 * p_f1
p_g_intersect_f2 <- p_g_on_f2 * p_f2
p_g <- p_g_intersect_f1 + p_g_intersect_f2
p_g
#c) p_f2_on_g
p_f2_on_g <- p_g_intersect_f2 / p_g
p_f2_on_g
#d) independent events?
p_f1 * p_g == p_g_intersect_f1
#e) p_f1_on_k (k is the union of u & h)
p_k_on_f1 <- p_u_on_f1 + p_h_on_f1
p_k_on_f2 <- p_u_on_f2 + p_h_on_f2
p_f1_on_k_nominator <- p_k_on_f1 * p_f1
p_f1_on_k_denominator <- p_k_on_f1 * p_f1 + p_k_on_f2 * p_f2
p_f1_on_k <- p_f1_on_k_nominator / p_f1_on_k_denominator
p_f1_on_k
```

## E7

s1: The email is spam | s2: The email is not spam
A: The email contains the word "Act"
B: The email contains the word "Buy"
C: The email contains the word "Call"

```{r}
#There are 25 spam email in 100 emails
p_s1 <- 0.25
p_s2 <- 0.75
#In 25 spam email, we have 20 with "Act", 15 with "Buy" and 10 with "Call"
p_a_on_s1 <- 20/25
p_b_on_s1 <- 15/25
p_c_on_s1 <- 10/25
#In 75 NOT_spam email, we have 3 with "Act", 6 with "Buy" and 9 with "Call"
p_a_on_s2 <- 3/75
p_b_on_s2 <- 6/75
p_c_on_s2 <- 9/75
#a) we are looking for: p_s1_on_a
p_s1_intersect_a <- p_a_on_s1 * p_s1
p_a <- p_a_on_s1 * p_s1 + p_a_on_s2 * p_s2
p_s1_on_a <- p_s1_intersect_a / p_a
p_s1_on_a
## Because p_s1_on_a < 0.9, this email should not be classified as spam
#b) we are looking for: p_s1_on_m (m is the intersection of a and c)
p_s2_intersect_a <- p_a_on_s2 * p_s2
p_a <- p_s1_intersect_a + p_s2_intersect_a
p_s1_intersect_c <- p_c_on_s1 * p_s1
p_s2_intersect_c <- p_c_on_s2 * p_s2
p_c <- p_s1_intersect_c + p_s2_intersect_c
p_a_intersect_c <- p_a * p_c # a & c are statistically independent
p_m <- p_a_intersect_c
p_s1_intersect_m <- p_m * p_s1
p_s1_on_m <- p_s1_intersect_m / p_m
p_s1_on_m
#c) we are looking for: p_s1_on_n (n is the intersection of a, b and c)
p_s1_intersect_b <- p_b_on_s1 * p_s1
p_s2_intersect_b <- p_b_on_s2 * p_s2
p_b <- p_s1_intersect_b + p_s2_intersect_b
p_a_intersect_b_intersect_c <- p_a * p_b * p_c
p_n <- p_a_intersect_b_intersect_c
p_s1_intersect_n <- p_n * p_s1
p_s1_on_n <- p_s1_intersect_n / p_n
p_s1_on_n
```

