---
title: "BE601 - Data Analytics I"
subtitle: "Seminar 2"
author: "Uyen Vu (uyen.vu@phdstudent.hhs.se)"
date: "13/08/2022"
output: word_document
monofont: 'consolas'
---

\newpage

# Test Exercise
3.17 - p.114
A department store manager has monitored the number of complaints received per week about poor service. The probabilities for numbers of complaints in a week, established by this review, are shown in the following table. Let A be the event “there will be at least one complaint in a week” and B the event “there will be fewer than ten complaints in a week.”

| Number of complaints |   0  | 1 to 3 | 4 to 6 | 7 to 9 | 10 to 12 | More than 12 |
|----------------------|:----:|:------:|:------:|:------:|:--------:|:------------:|
| Probability          | 0.14 |  0.39  |  0.23  |  0.15  |   0.06   |     0.03     |

a. Find the probability of A.
b. Find the probability of B.
c. Find the probability of the complement of A.
d. Find the probability of the union of A and B.
e. Find the probability of the intersection of A and B.
f. Are A and B mutually exclusive?
g. Are A and B collectively exhaustive?

```{r}
P_A <- 1 - 0.14
paste(c("a) P(A) is",P_A), collapse = " ")
P_B <- 1 - 0.06 - 0.03
paste(c("b) P(B) is", P_B), collapse = " ")
P_NotA <- 0.14
paste(c("c) P(NotA) is",P_NotA), collapse = " ")
P_A_intersect_B <- P_B - P_NotA
P_A_union_B <- P_A + P_B - P_A_intersect_B
paste(c("d) P(A_union_B) is",P_A_union_B), collapse = " ")
paste(c("e) P(A_intersect_B) is",P_A_intersect_B), collapse = " ")
P_A_intersect_B == 0
paste(c("f) Because P(A_intersect_B) is not equal to 0", "A and B are not mutually exclusive"), collapse = ", ")
P_A + P_B == 1
paste(c("g) Because P(A) + P(B) is not equal to 1", "A and B are not collectively exhaustive"), collapse = ", ")
```

3.18 - page 115
A corporation receives a particular part in shipments of 100. Research indicated the probabilities shown in the accompanying table for numbers of defective parts in a shipment.

| Number      |   0  |   1  |   2  |  3  | >3 defective |
|-------------|:----:|:----:|:----:|:---:|:------------:|
| Probability | 0.29 | 0.36 | 0.22 | 0.1 |     0.03     |

a. What is the probability that there will be fewer than three defective parts in a shipment?
b. What is the probability that there will be more than one defective part in a shipment?
c. The five probabilities in the table sum to 1. Why must this be so?

```{r}
E1 <- "E1: Fewer than 3 defective parts in a shipment"
E2 <- "E2: More than one defective part in a shipment"
E1
E2
P_E1 <- 0.29 + 0.36 + 0.22
P_E2 <- 0.29 + 0.36
paste(c("P(E1) = ", P_E1), collapse = " ")
paste(c("P(E2) = ", P_E2), collapse = " ")
```


# Discussion exercises
## 46 - page 125

A conference began at noon with two parallel sessions. The session on portfolio management was attended by 40% of the delegates, while the session on chartism was attended by 50%. The evening session consisted of a talk titled “Is the Random Walk Dead?” This was attended by 80% of all delegates.

a. If attendance at the portfolio management session and attendance at the chartism session are mutually exclusive, what is the probability that a randomly chosen delegate attended at least one of these sessions?
b. If attendance at the portfolio management session and attendance at the evening session are statistically independent, what is the probability that a randomly chosen delegate attended at least one of these sessions?
c. Of those attending the chartism session, 75% also attended the evening session. What is the probability that a randomly chosen delegate attended at least one of these two sessions?

```{r}
E1 <- "E1: Delegates attend portfolio management session"
E2 <- "E2: Delegates attend chartism session"
E3 <- "E3: Delegates attend evening session"
P_E1 <- 0.4
P_E2 <- 0.5
P_E3 <- 0.8
#a) assume that E1 & E2 are mutually exclusive
P_E1_union_E2 <- P_E1 + P_E2
paste(c("P(E1_union_E2) is", P_E1_union_E2), collapse = " ")
#b) assume that E1 & E3 are statistically independent
P_E1_intersect_E3 <- P_E1 * P_E3 #statistically independent
E4 <- "E4: Delegates attend atleast one of 2 sessions E1 & E3"
P_E4 <- P_E1 + P_E3 - P_E1_intersect_E3
paste(c("P(E4) is", P_E4), collapse = " ")

#c) conditional probability
P_E3_on_E2 <- 0.75
P_E3_intersect_E2 <- P_E3_on_E2 * P_E2 #multiplicative rule
E5 <- "E5: Delegates attend atleast one of 2 sessions E2 & E3"
P_E5 <- P_E2 + P_E3 - P_E3_intersect_E2
paste(c("P(E5) is", P_E5), collapse = " ")
```

## 51 - page 126

An editor may use all, some, or none of three possible strategies to enhance the sales of a book:

a. An expensive pre-publication promotion
b. An expensive cover design
c. A bonus for sales representatives who meet predetermined sales levels

In the past, these three strategies have been applied simultaneously to only 2% of the company’s books. Twenty percent of the books have had expensive cover designs, and, of these, 80% have had expensive pre-publication promotion. A rival editor learns that a new book is to have both an expensive pre-publication promotion and an expensive cover design and now wants to know how likely it is that a bonus scheme for sales representatives will be introduced. Compute the probability of interest to the rival editor.

```{r}
E1 <- "E1: Book has an expensive pre-publication promotion"
E2 <- "E2: Book has an expensive cover design"
E3 <- "E3: Book is associated with the bonus scheme for sales representatives"
E4 <- "E4: Book has both expensive pre-publication promotion and an expensive cover design"

P_E1_E2_E3_intersect <- 0.02
P_E2 <- 0.2
P_E1_on_E2 <- 0.8
P_E1_intersect_E2 <- P_E1_on_E2 * P_E2 #multiplication rule
P_E4 <- P_E1_intersect_E2
P_E3_intersect_E4 <- P_E1_E2_E3_intersect
P_E3_on_E4 <- P_E3_intersect_E4 / P_E4
paste(c("The probability of interest to the rival editor is", P_E3_on_E4), collapse = " ")
```

## E6

We define the events:

d1: item being defective

d2: item not being defective

c1: item being classified as defective

c2: item being classified as not_defective

```{r}
p_d1 <- 0.05
p_d2 <- 0.95
p_c2_on_d2 <- 0.98
p_c1_on_d1 <- 0.95
#a) Find p_c1
p_c1_intersect_d1 <- p_c1_on_d1 * p_d1
p_c2_intersect_d1 <- p_d1 - p_c1_intersect_d1
p_c2_intersect_d2 <- p_c2_on_d2 * p_d2
p_c1_intersect_d2 <- p_d2 - p_c2_intersect_d2
p_c1_intersect_d1
p_c2_intersect_d1
p_c2_intersect_d2
p_c1_intersect_d2
p_c1 <- p_c1_intersect_d1 + p_c1_intersect_d2
p_c1
#b) Find p_d1_on_c1
p_d1_on_c1 <- p_c1_intersect_d1 / p_c1
p_d1_on_c1
```

We can calculate the marginal probability of each event by using two-way table:

|            |       **D1**       |      **D2**      |      **Totals**      |
|------------|:------------------:|:----------------:|:--------------------:|
| **C1**     | 0.95*0.05=0.0475   | 0.02*0.95=0.019  | 0.0475+0.019=0.0665  |
| **C2**     | 0.05*0.05=0.0025   | 0.98*0.95=0.931  | 0.0025+0.0931=0.9335 |
| **Totals** | 0.0475+0.0025=0.05 | 0.019+0.931=0.95 |           1          |

#Additional recommended exercises

## 40
A mail-order firm considers three possible events in filling an order:
  A: The wrong item is sent.
  B: The item is lost in transit.
  C: The item is damaged in transit.
Assume that A is independent of both B and C and that B and C are mutually exclusive. The individual event probabilities are P(A) = 0.02, P(B) = 0.01, and P(C) = 0.04. Find the probability that at least one of these foul-ups occurs for a randomly chosen order.

```{r}
1- 0.96 * 0.99 * 0.98
```


## 3.99

In a campus restaurant it was found that 35% of all customers order vegetarian meals and that 50% of all customers are students. Further, 25% of all customers who are students order vegetarian meals.
a. What is the probability that a randomly chosen customer both is a student and orders a vegetarian meal?
b. If a randomly chosen customer orders a vegetarian meal, what is the probability that the customer is a student?
c. What is the probability that a randomly chosen customer both does not order a vegetarian meal and is not a student?
d. Are the events “customer orders a vegetarian meal” and “customer is a student” independent?
e. Are the events “customer orders a vegetarian meal” and “customer is a student” mutually exclusive?
f. Are the events “customer orders a vegetarian meal” and “customer is a student” collectively exhaustive?

S1: Customer is student | S1: Customer is not student
V1: Customer eats vegetarian | V2: Customer does not eat vegetarian

```{r}
#35% of all customers order vegetarian meals and 50% of all customers are students.
p_v1 <- 0.35
p_v2 <- 1 - p_v1
p_s1 <- 0.5
p_s2 <- 1 - p_s1
#25% of all customers who are students order vegetarian meals
p_v1_on_s1 <- 0.25
p_v2_on_s1 <- 1 - p_v1_on_s1
#a) randomly chosen customer both is a student and eats vegetarian: p_s1_intersect_v1
p_s1_intersect_v1 <- p_v1_on_s1 * p_s1
p_s1_intersect_v1
#b) p_s1_on_v1
p_s1_on_v1 <- p_s1_intersect_v1 / p_v1
p_s1_on_v1
#c) p_s2_intersect_v2 (we can draw a 2-way table)
p_s1_intersect_v2 <- p_v2_on_s1 * p_s1
p_s2_intersect_v2 <- p_v2 - p_s1_intersect_v2
p_s2_intersect_v2
#d) independent events?
p_v1 * p_s1 == p_s1_intersect_v1 #FALSE so not independent
#e) mutually exclusive? 
p_s1_intersect_v1 == 0 #FALSE so not mutually exclusive
#f) collectively exhaustive?
p_s1 + p_v1 == 1 #FALSE so not collectively exhaustive
```

## 3.102

A large corporation organized a ballot for all its workers on a new bonus plan. It was found that 65% of all night-shift workers favored the plan and that 40% of all female workers favored the plan. Also, 50% of all employees are night-shift workers and 30% of all employees are women. Finally, 20% of all night-shift workers are women.
a. What is the probability that a randomly chosen employee is a woman in favor of the plan?
b. What is the probability that a randomly chosen employee is either a woman or a night-shift worker (or both)?
c. Is employee gender independent of whether the night shift is worked?
d. What is the probability that a female employee is a night-shift worker?
e. If 50% of all male employees favor the plan, what is the probability that a randomly chosen employee both does not work the night shift and does not favor the plan?

Define the events:

G1: worker is female | G2: worker is male

N1: worker works night-shift | N2: worker does not work night-shift

F1: worker favors the plan | F2: worker does not favor the plan


```{r}
#65% of all night_shift workers favored the plan
p_f1_on_n1 <- 0.65
#40% of all female workers favored the plan
p_f1_on_g1 <- 0.4
#50% of all employees are night-shift workers
p_n1 <- 0.5
p_n2 <- 1 - p_n1 #complementary
#30% of all employees are women
p_g1 <- 0.3
p_g2 <- 1 - p_g1
#20% of all night-shift workers are women
p_g1_on_n1 <- 0.2
#a) a randomly chosen employee is a woman in favor of the plan: p_g1_intersect_f1
p_g1_intersect_f1 <- p_f1_on_g1 * p_g1 #multiplication rule
p_g1_intersect_f1
#b) a randomly chosen employee is either a woman or a nigh-shift worker (or both): p_g1_intersect_n1 & p_g1_union_n1
p_g1_intersect_n1 <- p_g1_on_n1 * p_n1
p_g1_intersect_n1
p_g1_union_n1 <- p_g1 + p_n1 - p_g1_intersect_n1
p_g1_union_n1
#c) employee gender independent of night shift is worked?
p_g1_intersect_n1 == p_g1 * p_n1 #not independent
#d) female employee is a night-shift worker: p_n1_on_g1
p_n1_on_g1 <- p_g1_intersect_n1/p_g1
p_n1_on_g1
#50% of all male employees favor the plan
p_f1_on_g2 <- 0.5
#e) employee both does work the night shift and does not favor the plan: p_n2_intersect_f2
p_g2_intersect_f1 <- p_f1_on_g2 * p_g2
p_f1 <- p_g1_intersect_f1 + p_g2_intersect_f1
p_n1_intersect_f1 <- p_f1_on_n1 * p_n1
p_n2_intersect_f1 <- p_f1 - p_n1_intersect_f1
p_n2_intersect_f2 <- p_n2 - p_n2_intersect_f1
p_n2_intersect_f2
```

## 3.101

f1: employee is female | f2: employee is male
g: employee had graduate degree | u: employee had undergraduate degree | h: employee had high school degree

```{r}
#80% of the employees are men & 20% of the employees are women
p_f1 <- 0.2
p_f2 <- 0.8
#men: 10% had graduate, 30% had undergraduate, 60% had high school
p_g_on_f2 <- 0.1
p_u_on_f2 <- 0.3
p_h_on_f2 <- 0.6
#women: 15% had graduate, 40% had undergraduate, 45% had high school
p_g_on_f1 <- 0.15
p_u_on_f1 <- 0.4
p_h_on_f1 <- 0.45
#a) p_f2_intersect_h
p_f2_intersect_h <- p_h_on_f2 * p_f2
p_f2_intersect_h
#b) p_g
p_g_intersect_f1 <- p_g_on_f1 * p_f1
p_g_intersect_f2 <- p_g_on_f2 * p_f2
p_g <- p_g_intersect_f1 + p_g_intersect_f2
p_g
#c) p_f2_on_g
p_f2_on_g <- p_g_intersect_f2 / p_g
p_f2_on_g
#d) independent events?
p_f1 * p_g == p_g_intersect_f1
#e) p_f1_on_k (k is the union of u & h)
p_k_on_f1 <- p_u_on_f1 + p_h_on_f1
p_k_on_f2 <- p_u_on_f2 + p_h_on_f2
p_f1_on_k_nominator <- p_k_on_f1 * p_f1
p_f1_on_k_denominator <- p_k_on_f1 * p_f1 + p_k_on_f2 * p_f2
p_f1_on_k <- p_f1_on_k_nominator / p_f1_on_k_denominator
p_f1_on_k
```

## E7

s1: The email is spam | s2: The email is not spam
A: The email contains the word "Act"
B: The email contains the word "Buy"
C: The email contains the word "Call"

```{r}
#There are 25 spam email in 100 emails
p_s1 <- 0.25
p_s2 <- 0.75
#In 25 spam email, we have 20 with "Act", 15 with "Buy" and 10 with "Call"
p_a_on_s1 <- 20/25
p_b_on_s1 <- 15/25
p_c_on_s1 <- 10/25
#In 75 NOT_spam email, we have 3 with "Act", 6 with "Buy" and 9 with "Call"
p_a_on_s2 <- 3/75
p_b_on_s2 <- 6/75
p_c_on_s2 <- 9/75
#a) we are looking for: p_s1_on_a
p_s1_intersect_a <- p_a_on_s1 * p_s1
p_a <- p_a_on_s1 * p_s1 + p_a_on_s2 * p_s2
p_s1_on_a <- p_s1_intersect_a / p_a
p_s1_on_a
## Because p_s1_on_a < 0.9, this email should not be classified as spam
#b) we are looking for: p_s1_on_m (m is the intersection of a and c)
p_s2_intersect_a <- p_a_on_s2 * p_s2
p_a <- p_s1_intersect_a + p_s2_intersect_a
p_s1_intersect_c <- p_c_on_s1 * p_s1
p_s2_intersect_c <- p_c_on_s2 * p_s2
p_c <- p_s1_intersect_c + p_s2_intersect_c
p_a_intersect_c <- p_a * p_c # a & c are statistically independent
p_m <- p_a_intersect_c
p_s1_intersect_m <- p_m * p_s1
p_s1_on_m <- p_s1_intersect_m / p_m
p_s1_on_m
#c) we are looking for: p_s1_on_n (n is the intersection of a, b and c)
p_s1_intersect_b <- p_b_on_s1 * p_s1
p_s2_intersect_b <- p_b_on_s2 * p_s2
p_b <- p_s1_intersect_b + p_s2_intersect_b
p_a_intersect_b_intersect_c <- p_a * p_b * p_c
p_n <- p_a_intersect_b_intersect_c
p_s1_intersect_n <- p_n * p_s1
p_s1_on_n <- p_s1_intersect_n / p_n
p_s1_on_n
```

