---
title: "BE601 - DATA ANALYTICS I"
subtitle: Chapter 3 - Probability - Book Exercise
author: "Uyen Vu (uyen.vu@phdstudent.hhs.se)"
date: "August 12th, 2022"
output: word_document
#    reference_docx: "assets/sse-word-template.docx"
#    toc: yes
#    toc_depth: 2

monofont: consolas
---


\newpage

# Example:

## 3.22 - page 132

A: be in a single accident
notA: not be in a single accident

B: had been drinking

```{r}
P_BonnotA <- 0.103
P_BonA <- 0.324
overinvolvement_ratio <- P_BonA / P_BonnotA
overinvolvement_ratio
```

Becaue the overinvolvement ration > 1, we can say that drinking increase the odds of being involved in car accidents.
0.06/0.87

# Exercise:

## 3.27

- define the outcome:

A1: use superior strategy			

A2: use normal strategy			

A3: use cheap strategy

B1: media ad

B2: price discount

B3: coupon

- event of interest: A1 በ B1

P(A1) = P(A2) = P(A3) = 1/3                     

P(B1) = P(B2) = P(B3) = 1/3

Because A & B are statistically independent:

P(A1 በ B1) = P(A) * P(B) = 1/9

## 3.28

```{r}
# number of ways to rank: 9!
factorial(9)
```

For each of the position, the chance to be correct is 1/9

## 3.29

E1: receive full scholarship

E2: receive 50% scholarship

E3: receive book

Possible outcomes for E1 = 100

Possible outcomes for E2 after E1 = 99

Possible outcomes for award = 100*99=9900

```{r}
# Permutation in R:
factorial(100)/factorial(98)
```

## 3.30

E1: Stock ranks 1st

E2: Stock ranks 2nd

E3: Stock ranks 3rd

Making the correction by chance is similar to choosing 3 stocks from the list of 6 and rank in order.

Possible permutations, with n = 6 and x = 3:

```{r}
factorial(6)/factorial(6-3)
```

## 3.31

G1: there is no graduate student

G2: there is at least 1 graduate student

U1: all are undergraduate student

G1 = U1

Possible outcomes for U1: combination with n = 4 and x = 3
All possible outcomes for choosing 3 out of 6: combination with n = 6 and x = 3

```{r}
# combination with n = 4 and x = 3
x1 <- factorial(4)/(factorial(3)*factorial(4-3))
# combination with n = 6 and x = 3
x2 <- factorial(6)/(factorial(3)*factorial(6-3))
# P(G1)
x1/x2
```

## 3.32

All possible outcomes (number of possible predictions): permutation with n = 7, x = 5

```{r}
n <- factorial(7)/factorial(7-5)
```

In all possible predictions, one will be correct. Probability of making the correct prediction:

```{r}
na <- 1/n
na
```

## 3.33

This is the combination with n = 8, x = 5

```{r}
factorial(8)/(factorial(5)*factorial(8-5))
```

## 3.34

a.  This is the combination with n = 7, x = 5

```{r}
all <- factorial(7)/(factorial(5)*factorial(7-5))
all
```

b. Possibilities that I am NOT chosen is combination with n = 6, x = 2

```{r}
# All possible outcomes
all <- factorial(7)/(factorial(5)*factorial(7-5))
all
# Possibilities that I am NOT chosen
i_am_not_chosen <- factorial(6)/(factorial(2)*factorial(6-2))
i_am_not_chosen
# Possibilities that I am chosen
i_am_chosen <- all - i_am_not_chosen
i_am_chosen
#possibilities that I am chosen to play heroine
i_am_chosen_heroine <- i_am_chosen * 0.5
i_am_chosen_heroine
```

c. Possibilities for me to play best friend:

```{r}
# Probability for the chosen person to play best friend is 0.5
P_best_friend <- 0.5
# All possible outcomes
all <- factorial(7)/(factorial(5)*factorial(7-5))
all
# Possibilities that I am NOT chosen
i_am_not_chosen <- factorial(6)/(factorial(2)*factorial(6-2))
i_am_not_chosen
# Possibilities that I am chosen
i_am_chosen <- all - i_am_not_chosen
i_am_chosen
# Possibilities that I am chosen to play best friend
i_am_chosen_best_friend <- i_am_chosen*0.5
i_am_chosen_best_friend
```

d. From a. & b. 

```{r}
# Probability that I am chosen to play heroine (using results from a & b)
P_i_am_chosen_heroine <- 3/21
P_i_am_chosen_heroine
```

Alternatively,

A: I am chosen

B: Playing heroine (P(B|A) = 0.5)

```{r}
# P(notA) is combination with n = 6, x = 2 
notA <- factorial(6)/(factorial(2)*factorial(6-2))
notA
# all possible outcome is combination with n = 7, x = 2
all <- factorial(7)/(factorial(2)*factorial(7-2))
all
# P(A) = 1 - P(notA)
P_A <- 1 - notA/all
P_A
# P(A በ B) = P(B|A) * P(A)
P_AintersectB <- 0.5 * P_A
P_AintersectB
```

e. Possibilities for me to play one of the 2 parts: 3 + 3 = 6

All possible outcomes: 21

```{r}
# Probability that I am chosen:
6/21
```

Alternatively,

```{r}
# P(notA) is combination with n = 6, x = 2 
notA <- factorial(6)/(factorial(2)*factorial(6-2))
# all possible outcome is combination with n = 7, x = 2
all <- factorial(7)/(factorial(2)*factorial(7-2))
# P(A) = 1 - P(notA)
P_A <- 1 - notA/all
P_A
```

## 3.25

a. Choose pilot: combination with n = 10, x = 2

Choose flight attendants: combination with n = 20, x = 5

```{r}
# Ways to choose pilots:
pilot <- factorial(10)/(factorial(2)*factorial(10-2))
pilot
#ways to choose attendants:
attendant <- factorial(20)/(factorial(5)*factorial(20-5))
attendant
#way to choose 2 pilots and 5 attendants:
crew <- pilot * attendant
crew
```

b.Possibilities for the sibling not being chosen from 20 attendants is combination with n = 19, x = 4



```{r}
## possibilities for the sibling NOT being chosen from 20 attendants is:
NOT_sibling_attendant <- factorial(19)/(factorial(5)*factorial(19-5))
NOT_sibling_attendant
## All possible outcomes for choosing attendants:
all_attendant <- factorial(20)/(factorial(5)*factorial(20-5))
all_attendant
## Probability that the sibling being chosen from 20 attendants is:
P_sibling_attendant <- 1 - NOT_sibling_attendant/all_attendant
P_sibling_attendant

## possibilities for the sibling NOT being chosen from 10 pilots is:
NOT_sibling_pilot <- factorial(9)/(factorial(2)*factorial(9-2))
NOT_sibling_pilot
## All possible outcomes for choosing pilots:
all_pilot <- factorial(10)/(factorial(2)*factorial(10-2))
all_pilot
## Probability that the sibling being chosen from 10 pilots is:
P_sibling_pilot <- 1 - NOT_sibling_pilot/all_pilot
P_sibling_pilot
## Probability for both siblings being chosen (muliplication rule):
P_both_sibling <- P_sibling_attendant * P_sibling_pilot
P_both_sibling
```

c. A: sibling is chosen as pilot

B: sibling is chosen as attendant

P(A U B) = P(A) + P(B) - P(A በ B)
Possibilities that neither of siblings is chosen: 1 - P(A U B)

```{r}
## possibilities for the sibling being chosen as attendant:
sibling_attendant <- factorial(19)/(factorial(4)*factorial(19-4))
sibling_attendant
## All possible outcomes for choosing attendants:
all_attendant <- factorial(20)/(factorial(5)*factorial(20-5))
all_attendant
## Probability that the sibling being chosen from 20 attendants is (P(A)):
P_sibling_attendant <- sibling_attendant/all_attendant
P_sibling_attendant

## possibilities for the sibling being chosen as pilot:
sibling_pilot <- factorial(9)/(factorial(1)*factorial(9-1))
sibling_pilot
## All possible outcomes for choosing pilots:
all_pilot <- factorial(10)/(factorial(2)*factorial(10-2))
all_pilot
## Probability that the sibling being chosen from 10 pilots is (P(B)):
P_sibling_pilot <- sibling_pilot/all_pilot
P_sibling_pilot
## Probability for both siblings being chosen (muliplication rule) (P(A በ B)):
P_both_sibling <- P_sibling_attendant * P_sibling_pilot
P_both_sibling
## Probability for at least 1 sibling is chosen:
P_atleast_one_sibling <- P_sibling_attendant + P_sibling_pilot - P_both_sibling
## Probability that no sibling is chosen:
P_neither_sibline <- 1 - P_atleast_one_sibling
P_neither_sibline
```

## 3.36

a. 

```{r}
# Choices for 2 private schools: combination (n=5,x=2)
private <- factorial(5)/(factorial(2)*factorial(5-2))
private
# Choices for 2 public schools: combination (n=3,x=2)
public <- factorial(3)/(factorial(2)*factorial(3-2))
public
# Possible choices: 
all <- private * public
all
```

b.

```{r}
# Choices that no private school is not underperformed (n=4,x=2)
private_not_underperformed <- factorial(4)/(factorial(2)*factorial(4-2))
private_not_underperformed
#Choices that no public school is not underperformed (n=2,x=2)
public_not_underperformed <- factorial(2)/(factorial(2)*factorial(2-2))
public_not_underperformed
#Choices that neither sector has underperformed school:
not_underperformed <- private_not_underperformed * public_not_underperformed
not_underperformed
# All possible choices
all <- 30
# Probability that at least one of the schools will underperform:
P_atleast_underperform <- 1 - not_underperformed/all
P_atleast_underperform
```

## 3.37

E1: being concerned about employment prospect - P(E1) = 0.6

E2: being concerned about grades - P(E2) = 0.45

P(E1 በ E2) = 0.3


```{r}
P_E1 <- 0.6
P_E2 <- 0.45
P_intersection <- 0.3
P_union <- P_E1 + P_E2 - P_intersection
P_union
```

## 3.38

A: ask for help

E: making a purchase

```{r}
P_A <- 0.3
P_E <- 0.2
P_intersection <- 0.15
P_union <- P_A + P_E - P_intersection
P_union
```

## 3.39

A: immediate donation - P(A) = 0.75

B: request information - P(B) = 0.15

C: no interest = 0.3

D: donation | B - P(D) = 0.2

E: no donation | B - P_notD = 0.8

a)

```{r}
P_A <- 0.75
P_notA <- 1 - P_A
# The probability of having no immediate donation after 4 calls:
P_notA_4calls <- P_notA^4
P_notA_4calls
```

b)

```{r}
#Probability that each call is unsuccessful
P_B <- 0.15
P_C <- 0.3
P_notD <- 0.8
P_fail <- P_C + P_B * P_notD
#Probability that 4 calls is unsuccessful in a row
P_4successful <- P_fail^4
P_4successful
```

## 3.40

P(B በ C) = 0 (mutually exclusive)

P(A በ B በ C) = 0

P(A በ B) = P(A)*P(B) (statistically independent)

P(A በ C) = P(A)*P(C) (statistically independent)

<Draw the vein diagram here>

P(A U B U C) = P(A) + P(B) + P(C) - P(A በ B) - P(A በ C) - P(B በ C) - P(A በ B በ C)

```{r}
P_A <- 0.02
P_B <- 0.01
P_C <- 0.04
P_unionABC <- P_A + P_B + P_C - P_A*P_B - P_A*P_C - 0 - 0
P_unionABC
```

## 3.41

0.98 * 0.85

## 3.42

0.1/0.18
0.1/0.12

## 3.43

0.01/0.08




